---
title: "Comparing the performance of Purple Air sensors to DEQ reference nephelometers"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
source("purpleair_utility.R")

# read in the Purple Air data.  
sensors <- read_sensor_data()



# specify the reference monitor data files and names 
ref_files <- c("data/deq_stns/SEL 5-1-2017 - 9-5-2018 pm25 1 min.csv",
               "data/deq_stns/TBC 5-1-2017 - 9-5-2018 pm25 1min.csv",
               "data/deq_stns/BHP 5-1-2017 - 9-5-2018 pm25 1min clean.csv",
               "data/deq_stns/HHF 5-1-2017 - 9-5-2018 pm25 1 min.csv")
ref_name <- c("SEL", "TBC", "BHP", "HHF")
nephs <- read_ref_monitor_data(monitor_file_list = ref_files, monitor_name_list = ref_name)

sensors_1h <- create_1hour_df(sensors)
nephs_1h <- create_1hour_df(nephs)
sen_stat_1h <- stats_summ(sensors_1h, "pm25")
ref_stat_1h <- stats_summ(nephs_1h, "neph")

dist_df <- create_distance_matrix()
dist_long <- gather(dist_df[, c(4:7,1)], key = ref, value = dist, -site)
dist_long$ref <- substring(dist_long$ref, 5, 7)
corr_df <- corr_summ(sensors_1h, nephs_1h)
corr_long <- gather(corr_df[, c(2:5,1)], key = ref, value = corr, -site)
corr_long$ref <- substring(corr_long$ref, 5, 7)

corr_dist <- merge(corr_long, dist_long, by = c("site", "ref"))
corr_dist$group <- substring(corr_dist$site, 4,5)
corr_dist$sensor <- substring(corr_dist$site, 1,3)

```

This documents compares the PM2.5 readings of ~14 Purple Air sensors in the Portland Metro area with the scattering (bscat) measured by 4 Oregon DEQ nephelometers in the area.

#### Summary statistics for the reference monitors
```{r ref_stat, echo = FALSE, results = 'asis'}
library(flextable)
library(officer)
tb <- regulartable(ref_stat_1h, col_keys = c("Count", "First Obs.", "Last Obs.", "Min",
                                             "Max", "Mean", "Median", "Std. Dev."))
knitr::knit_print(tb)
```

#### Summary statistics for the reference monitors
```{r sen_stat, echo = FALSE}
knitr::kable(sen_stat_1h)
```

#### Correlation between the Purple Air sensors and reference nephs
```{r correl, echo=FALSE}
knitr::kable(corr_df)
```

#### Graph showing correlation with distance, colored by Purple Air sensor
```{r dist-corr-by-sensor, echo = FALSE}
library(ggplot2)
g <- ggplot(data = corr_dist, aes(dist, corr )) + 
  theme_minimal() +
  geom_point(aes(color = sensor, fill = sensor, shape = group, size = group)) +
  scale_shape_manual(values=c(1, 19)) + 
  scale_size_manual(values = c(5.5, 4.5)) +
  scale_color_manual(values = c("#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072", "#80B1D3", 
                                "#FDB462", "#B3DE69", "#FCCDE5", "#D9D9D9", "#BC80BD",
                                "#CCEBC5", "#FFED6F", "#4DAF4A", "#984EA3"
                                )) +
  guides(shape = guide_legend(title.position = "top", ncol = 1)) +
  guides(color = guide_legend(title.position = "top", nrow = 2)) +
  theme(legend.position = "bottom") +
  xlab("Distance from reference neph") +  ylab("Pearson correlation") +
  ggtitle("Correlation of Purple Airs with reference nephelometers")
plot(g)
```

#### Graph showing correlation with distance, colored by reference neph
```{r dist-corr-by-ref, echo = FALSE}
#library(ggplot2)
r <- ggplot(data = corr_dist, aes(dist, corr )) + 
  theme_minimal() +
  geom_point(aes(fill = ref, color = ref, shape = group, size = group)) +
  scale_shape_manual(values=c(1, 19)) + 
  scale_size_manual(values = c(5.5, 4.5)) +
  scale_color_manual(values = c("#FB8072", "#80B1D3", "#FDB462", "#B3DE69" )) +
  guides(shape = guide_legend(title.position = "top", nrow = 1)) +
  guides(color = guide_legend(title.position = "top", nrow = 1)) +
  theme(legend.position = "bottom") +
  xlab("Distance from reference neph") +  ylab("Pearson correlation") +
  ggtitle("Correlation of Purple Airs with reference nephelometers")
plot(r)

```

